---
title: "ST5226 Spatial Statistics - Areal (Regional) Data"
author: "Yeo Ming Jie, Jonathan (A0164616A)"
output: 
  html_document:
    toc: true
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) # clear global directory
knitr::opts_knit$set(root.dir = '/Users/jyeo_/Desktop/MSc Statistics Coursework/ST5226 Spatial Statistics/Data')
```

Loading Required Packages:

```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(sp)
library(spdep)
library(rgdal)
library(RColorBrewer)
library(classInt)
library(gstat)
library(nlme)       # For Generalised Least Squares Estimation
library(SpatialEpi) # For Kulldorff scan statistic 
```

## Lecture 11: Areal Data

Loading regional data for Leukemia cases in New York States:

```{r plot_Syracuse, warning = FALSE}
NY8 <- readOGR("NY_data", "NY8_utm18", verbose=FALSE)
Syracuse <- NY8[NY8$AREANAME == "Syracuse city", ] # SpatialPolygonsDataFrame
plot(Syracuse, axes=TRUE, main="Syracuse City Census Tracts", 
     cex.axis=0.8)
```

### Defining Neighbours using K-Nearest Neighbours

The `knearneigh` function returns a matrix where the $i$th row gives the indices of the $k$-nearest neighbours of region $i$.

```{r neighbours_1}
coords <- coordinates(Syracuse)
IDs <- row.names(slot(Syracuse, "data"))
# Consider only the nearest region to be a neighbour
Sy8_nb <- knn2nb(knearneigh(coords, k=1), row.names = IDs)
# Consider the 2 nearest regions to be neighbours
Sy9_nb <- knn2nb(knearneigh(coords, k=2), row.names = IDs)
Sy8_nb # knn2nb converts matrix to an `nb` object (list)
```

Plotting neighbours:

```{r neighbours_2}
opar <- par(mfrow=c(1,2))
# 1-Nearest Neighbour Links plotted
plot(Syracuse, axes=TRUE, cex.axis=0.8)
plot(Sy8_nb, coords, col='red', add=TRUE) 

# 2-Nearest Neighbour Links plotted
plot(Syracuse, axes=TRUE, cex.axis=0.8)
plot(Sy9_nb, coords, col='red', add=TRUE)
par(opar)
```

```{r neighbours_3}
# Identifying number of sub-graphs
n.comp.nb(Sy8_nb)$nc
```

### Distance-based Neighborhood Definitions

```{r neighbours_4}
max_1nn <- max(unlist(nbdists(Sy8_nb, coords)))
max_1nn # identifying upper bound
```

```{r neighbours_5}
# Regions within a distance 0.75*1544 considered neighbours.
Sy10_nb <- dnearneigh(coords, d1=0, d2=0.75*max_1nn, row.names = IDs)
# Regions within a distance 1.5*1544 considered neighbours.
Sy11_nb <- dnearneigh(coords, d1=0, d2=1.5*max_1nn, row.names = IDs)
summary(Sy10_nb) # Notice that 2 regions have no neighbours.
```
```{r}
# Count the number of neighbour links:
sapply(list(Sy10_nb, Sy11_nb), function(x) sum(card(x)))
```

### Adjacency-based neighbourhood definitions

Initialize rook-style neighborhood structure - regions intersect along a line. For queen style, regions intersect at either line or point:

```{r neighbours_6}
# Create rook-style contiguity structure.
NY_nb <- poly2nb(NY8, queen = FALSE)
summary(NY_nb)
n.comp.nb(NY_nb)$nc # Number of disconnected subgraphs
```

### Assigning Spatial Weights

Note that there are two types of spatial weights that can be assigned: $W$-style (average) weights and $B$-style (binary) weights.

```{r weights_1}
Sy9_nb_W <- nb2listw(Sy9_nb)
Sy9_nb[[1]] # note weights sum to unity
Sy9_nb_W$weights[[1]]
```

Binary-style weights:

```{r weights_2}
Sy9_nb_B <- nb2listw(Sy9_nb, style="B") # Binary-style weights
Sy9_nb_B$weights[[1]]
Sy9_nb_B$weights[[2]]
```

### Measures of Spatial Autocorrelation: Moran's I 

```{r moran_1}
# Create binary weights matrix
wts <- nb2listw(NY_nb, style='B')
# Moran's I test under Normality assumption
moran.test(NY8$TRACTCAS, wts, randomisation = FALSE)
```

```{r moran_2}
moran.mc(NY8$TRACTCAS, wts, nsim=999)
```

```{r moran_3}
# A function to compute the p-value of Moran's I test statistic under 
# the constant risk hypothesis.
moran.pois <- function(y, n_vec, listw, nsim) {
  Tstat <- rep(0, nsim)
  # Compute Moran's I from observed data
  Tstat[1] <- moran(y, listw, length(y), Szero(listw))$I
  
  cr <- sum(y)/sum(n_vec) # estimate constant risk
  pmeans <- cr*n_vec      # expected values for each region
  for(ii in 2:nsim) {
    # Generate from expected values (Poisson)
    tmp <- rpois(rep(1, length(pmeans)), pmeans)
    # Compute Moran's I for simualted data set 
    Tstat[ii] <- moran(tmp, listw, length(y), Szero(listw))$I
  }
  
  sum(Tstat[-1] > Tstat[1])/(nsim-1)
}
set.seed(111)
moran.pois(NY8$TRACTCAS, NY8$POP8, wts, 999)
```

### Moran Coefficient Plot - Evaluating Spatial Autocorrelation

We first standardize our dataset.

```{r moran_4}
cases <- NY8$TRACTCAS          # Observed values
popn <- NY8$POP8               # population values
cr <- sum(cases)/sum(popn)     # constant risk
Ei <- cr * NY8$POP8            # expected values
Ystd <- (cases - Ei)/sqrt(Ei)  # standardised values
moran.plot(Ystd, wts)          # create moran plot
```

* The plot indicates that there are several influential points in the cloud.
* We suspect that, if any, there is only weak spatial autocorrelation.
* Remember that we have not incorporated spatial covariates yet. When we do so, we would make the Moran coefficient plot with the residuals to assess if there is any residual spatial autocorrelation.

### Geary’s c

```{r geary_1}
geary.test(NY8$TRACTCAS, wts, randomisation = FALSE)
geary.mc(NY8$TRACTCAS, wts, nsim=9999)
```

```{r geary_2}
geary.pois <- function(y, n_vec, listw, nsim) {
  Tstat <- rep(0, nsim)
  Tstat[1] <- geary(y, listw, length(y), length(y)-1, Szero(listw))$C
  
  cr <- sum(y)/sum(n_vec)
  pmeans <- cr*n_vec
  for(ii in 2:nsim) {
    tmp <- rpois(rep(1, length(pmeans)), pmeans)
    Tstat[ii] <- geary(tmp, listw, length(y), length(y)-1, Szero(listw))$C
  }
  
  sum(Tstat[-1] < Tstat[1])/(nsim-1)
}
set.seed(111)
geary.pois(NY8$TRACTCAS, NY8$POP8, wts, 999)
```

### Local Moran's I (Measure of local spatial association)

```{r local_moran}
pal <- brewer.pal(5, 'Reds')
tmp <- localmoran(NY8$TRACTCAS, wts) 
NY8$lM <- tmp[,4] # Extract standardised version of local I
q5 <- classIntervals(tmp[,4], n=5, style='fisher')
q5$brks[6] <- 11.6
spplot(NY8, 'lM', col.regions=pal, at=q5$brks, main="Local Moran's I")
```

## Lecture 12: Spatial Regression for Regional Data

```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(sp)
library(spdep)
library(rgdal)
library(RColorBrewer)
library(classInt)
library(gstat)
library(nlme)       
library(SpatialEpi) 
library(spatialreg)
```


```{r read_NY8, warning=FALSE}
NY8 <- readOGR("NY_data", "NY8_utm18", verbose=FALSE)
```

```{r NY8_plots}
opar <- par(mfrow=c(2,2))
props <- NY8$TRACTCAS/NY8$POP8
hist(props, main="Raw Proportions")
qqnorm(props)
qqline(props)
hist(NY8$Z, main="Transformed Proportions")
qqnorm(NY8$Z)
qqline(NY8$Z)
par(opar)
```

```{r NY8_OLS}
NY_coords <- as.data.frame(coordinates(NY8))
names(NY_coords) <- c("x", "y")
# Create a data frame that includes the x and y coordinates as well.
NY_data <- cbind(slot(NY8, "data"), NY_coords)
# OLS analysis of NY data
NY_ols <- gls(Z ~ PEXPOSURE + PCTOWNHOME + PCTAGE65P, 
              data=NY_data, method="ML")
summary(NY_ols)
```

```{r plot_NY8_OLS}
plot(NY_ols)
```

```{r NY8_wls}
# WLS analysis of NY data
NY_wls <- gls(Z ~ PEXPOSURE + PCTOWNHOME + PCTAGE65P, 
              data=NY_data, method='ML',
              weights=varFixed(~1/POP8))
summary(NY_wls)
```

```{r plot_NY8_wls, warning=FALSE}
pal1 <- brewer.pal(n=6, 'RdBu')
NY8$res1 <- residuals(NY_wls, type='pearson')
q6 <- classIntervals(NY8$res1, 6, style='fisher')
q6$brks[6] <- 2.1311
spplot(NY8, 'res1', at=q6$brks, col.regions=pal1, main='Residuals Chloropleth')
```

```{r moran_plot}
NY_nb <- poly2nb(NY8, queen = FALSE)
wts <- nb2listw(NY_nb, style='B')
moran.plot(as.vector(NY8$res1), wts)
```

```{r NY8_resplot}
opar <- par(mfrow=c(1,2))
hist(NY8$res1)
qqnorm(NY8$res1)
qqline(NY8$res1)
par(opar)
```

```{r res_svgm}
res_svgm <- variogram(res1 ~ 1, NY8)
plot(res_svgm)
```

```{r NY_gls}
NY_gls <- gls(Z ~ PEXPOSURE + PCTOWNHOME + PCTAGE65P, data=NY_data,
              weights=varFixed(~1/POP8), method='ML',
              corr = corSpher(value=c(15000, 0.8), form=~x+y, nugget=TRUE))
summary(NY_gls)
```

```{r NY_anova}
# Compare the ols, wls and gls models, using hypothesis testing 
# where possible, and using AIC where models are not nested.
anova(NY_ols, NY_wls, NY_gls)
```

```{r NY_sar}
nysar <- spautolm(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME, 
                  data=NY8, listw=wts)
summary(nysar)
```

```{r NY_sar2}
nysar2 <- spautolm(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME, data=NY8, listw=wts, 
                   weights=POP8)
summary(nysar2)
```

```{r NY_car}
nycar <- spautolm(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME, data=NY8, listw=wts, 
                  family="CAR")
nycar2 <- spautolm(Z ~ PEXPOSURE + PCTAGE65P + PCTOWNHOME, data=NY8, listw=wts, 
                  family="CAR", weights=POP8)
summary(nycar)
summary(nycar2)
```








