---
title: "ST5226 Spatial Statistics - Geostatistics and Spatial Predictions"
author: "Yeo Ming Jie, Jonathan (A0164616A)"
output: 
  html_document:
    toc: true
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) # clear global directory
knitr::opts_knit$set(root.dir = '/Users/jyeo_/Desktop/MSc Statistics Coursework/ST5226 Spatial Statistics/Data')
```

## Lecture 7: Introduction to Geostatistics

Loading required packages:

```{r, message=FALSE, warning = FALSE}
library(sp)
library(gstat)
library(rgdal)
library(RColorBrewer)
library(classInt)
```

### Meuse Dataset

Recall the Meuse river bank dataset which contained measurements of the metal concentration along the Meuse river in the Netherlands.

```{r Meuse River Log-Zn}
data(meuse)
coordinates(meuse) <- ~x+y # converts meuse into SpatialPointsDataFrame
meuse$lzn <- log(meuse$zinc)
spplot(meuse, "lzn", main="Log-Zinc Measurements", 
       col.regions = heat.colors(5))
```

Observe that the zinc concentrations is larger close to the river bank, clear spatial trend in a given direction - zinc concentrations decreasing as we move in a direction perpendicular to the river bank.

### Exploratory Variogram Analysis

Checking for existence/strength of spatial correlation from scatter plots of pairs $Z(s_i)$ and $Z(s_j)$ grouped according to their separation distances $h = ||s_i-s_j||$, i.e. h-scatter plots. Consider the plots for the following sets of distance bins:

```{r h scatterplot}
hscat(log(zinc) ~ 1, meuse, (0:9)*100)
```

Plots indicate the correlation coefficient $r$, observe correlation decays with increasing distance. 

Next, we plot the sample semivariogram estimator for meuse. Observe the following key features, e.g. nugget, sill $\approx 0.6$, Range $\approx 800$.

```{r, Sample Semivariogram Meuse}
vgm_log_zinc <- variogram(log(zinc) ~ 1, meuse) # Default in the North-South direction
# ~ 1 defines a single constant predictor, leading to a spatially constant mean coefficient
plot(vgm_log_zinc)
```

We can also investigate anisotropy i.e. the covariance function $C(\cdot)$ of the process is a function of both distance and direction, by computing the sample variogram in 4 different directions.

```{r Anisotropy}
# alpha vector of angles in degrees, clockwise from North
plot(variogram(log(zinc) ~ 1, meuse, alpha=c(0, 45, 90, 135)))
```

Observe that the variogram grows slowly in the NE direction, but grows most rapidly in the SE direction.

### Smoky Mountains Dataset (pH measurements)

Next, consider the dataset containing the pH measurements at locations within the Great Smoky Mountains. Observe from the plot that observations with high pH values lay around the boundary of the region.

```{r Smoky Mountains Data, message = F}
smoky_data <- read.table("/Users/jyeo_/Desktop/MSc Statistics Coursework/ST5226 Spatial Statistics/Data/smoky/smoky.dat", header=TRUE, skip=1)
coordinates(smoky_data) <- ~easting + northing # initialize Spatial Points Data Frame
smok <- readOGR("smoky", "smokypoly") # Spatial Polygons
pal <- brewer.pal(4, "RdYlGn")
q4 <- classIntervals(smoky_data$ph, n=4, style="fisher")
spplot(smoky_data, "ph", cuts=q4$brks , col.regions=pal, colorkey=TRUE, 
       sp.layout=list("sp.polygons", smok), scales=list(draw=TRUE))
```

Note that the variogram object contains information on the exact values of $h$ used, along with the nuymber of points used to compute the estimate. Plotting the sample variogram estimator, we obtain the following:

```{r Smoky Mountain Semivariogram, collapse = T}
vgm_ph <- variogram(ph ~ 1, smoky_data, cutoff=120, width=10)
head(vgm_ph, n=2)
plot(vgm_ph)
```

### Available Semivariogram Types

```{r}
show.vgms(par.strip.text=list(cex=0.6))
```

In order to get a full list of variogram names used in `gstat`, we can type `vgm()`. The `vgm()` function is used to generate variogram models (or add to existing models).

```{r}
head(vgm())
```

### Parametric Semivariograms

Suppose we wish to create a Spherical variogram with parameters $c_0=1$ (nugget), $c_s=2$ (partial sill) and $a_s = 10$ (range).

```{r Fit Spherical Semivariogram}
sph1 <- vgm(psill=2, model='Sph', range=10, nugget=1)
class(sph1); sph1
# Plot of a particular semivariogram (Spherical)
show.vgms(sill=2, range=10, models='Sph', nugget=1, max=13)
```

Alternatively, consider the Gaussian semivariogram model with effective range $3a_e = 10$

```{r}
gauss1 <- vgm(psill=2, model='Gau', range=10, nugget=1)
gauss1
```

### Adding Semivariograms

The sum of two valid semivariograms is also a valid semivariogram. Consider the addition of the Spherical and Gaussian semivariograms.

```{r Add Semivariograms}
gauss2 <- vgm(psill=2, model='Gau', range=10, nugget=0, add.to=sph1)
out <- variogramLine(gauss2, maxdist=15)
plot(out, type='l', main="Sum of Two Variograms")
```

### Fitting Semivariograms using Likelihood-based Methods

```{r Fitting Spherical Semivariogram to Data}
fit_v <- fit.variogram(vgm_log_zinc, vgm(1, "Sph", 800, 1))
fit_v
attr(fit_v, "SSErr")
plot(vgm_log_zinc, model=fit_v)
```

## Lecture 8: Spatial Prediction (Kriging)

```{r, Meuse IDW}
data(meuse.grid)
coordinates(meuse.grid) <- ~x+y
meuse.grid <- as(meuse.grid, "SpatialPixelsDataFrame")

idw.out <- gstat::idw(log(zinc) ~ 1, meuse, meuse.grid, idp=2.5)
plot(idw.out, col=heat.colors(100))
head(as.data.frame(idw.out), n=2)
```


```{r Original Plot, warning=FALSE, message=FALSE, fig.height=6}
spplot(meuse, "lzn", main="Log-Zinc Measurements", 
       col.regions = heat.colors(4),key.space="right")
```


```{r Kriging}
vgm_log_zinc <- variogram(log(zinc) ~ 1, meuse)
fit_v <- fit.variogram(vgm_log_zinc, vgm(1, "Sph", 800, 1))
lz.ok <- krige(log(zinc)~1, meuse, meuse.grid, fit_v)
plot(lz.ok, col=heat.colors(100))
```

```{r Kriging Variance Plot}
plot(lz.ok['var1.var'])
plot(meuse, add=TRUE, col='grey50')
```


```{r Filtering}
err.fit <- fit.variogram(vgm_log_zinc, vgm(1, "Sph", 800, Err=1))
krige(log(zinc) ~ 1, meuse, meuse[1,], err.fit)
krige(log(zinc) ~ 1, meuse, meuse[1,], fit_v)
```


```{r Universal Kriging}
vt.fit2 <- fit.variogram.gls(log(zinc) ~ sqrt(dist), data=meuse[1:40,], 
                             model= vgm(1, "Exp", 300, 1) ,trace=FALSE)
lz.uk.gls <- krige(log(zinc)~sqrt(dist), meuse, meuse.grid, vt.fit2)
plot(lz.uk.gls, col=heat.colors(100))
```


## Lecture 9: Spatial Random Effects Model (Meuse data)

Required Packages:

```{r load_libraries, message = FALSE, warning=FALSE}
library("sp")
library("FRK")
library("gstat")
library("ggplot2")
library("gridExtra")
library("grid")
#library("INLA")
```

### Construcing Basic Areal Units (BAUs)

```{r create_BAUs}
data(meuse)
coordinates(meuse) = ~x+y # changes meuse into SpatialPointsDataFrame

# BAUs from hexagonal grid
HexPols_df <- auto_BAUs(manifold = plane(),
                        cellsize = 100,
                        type = "hex",
                        data = meuse,
                        nonconvex_hull = FALSE) 
# BAUs cover the convex hull of observation locations
plot(HexPols_df)
```

### Constructing the basis functions

```{r create_basis}
# Constructing basis functions
G <- auto_basis(manifold = plane(),data=meuse, # manifold - nature of spatial domain, plane() or sphere()
                nres = 2,regular=2,prune=0.1,type = "bisquare") 
# nres - number of resolutions (in this case, small and large scale, 2 resolutions)

show_basis(G,ggplot()) + geom_point(data=data.frame(meuse),aes(x,y))
```

### Fitting the Spatial Random Effects (SRE) Model

```{r fit_SRE}
set.seed(1)

# formula for SRE model (just an intercept here)
f <- log(zinc) ~ 1

# Call FRK()
S <- FRK(f = f,                # formula
         data = list(meuse),   # list of data objects (just meuse in this case)
         basis=G,
         BAUs=HexPols_df)          

summary(S)   # Print out a summary of the returned (fitted) SRE model
```

### Prediction

```{r prediction}
# Get predictions (for the BAUs by default)
GridBAUsPred<- predict(S, obs_fs = FALSE)

# Convert the SpatialPolygonsDataFrame "GridBAUs1" to an ordinary data frame
# For plotting:
Pred_df <- SpatialPolygonsDataFrame_to_df(sp_polys = GridBAUsPred,vars=c("mu","sd"))

# Set coordinate reference system for predictions
# 4326 is the EPSG code for WGS84

# Create plot for prediction means, mu
g1<-ggplot() +
  geom_polygon(
    aes(x=x, y=y, fill=mu, group=id), # fill color by mu
    data = Pred_df)

# Change x and y axis labels
g1<-g1 + xlab("Easting") + ylab("Northing")

# Plot for standard deviation
g2<-ggplot() +
  geom_polygon(
    aes(x=x, y=y, fill=sd, group=id),
    data = Pred_df)

g2<-g2 + xlab("Easting") + ylab("Northing")

# Combine the above two plots on one graph
grid.arrange(g1,g2,nrow=1)
```

## Lecture 10:

```{r load_packages}
library("sp")
library("FRK")
library("gstat")
library("ggplot2")
library("gridExtra")
library("grid")
# library("INLA")
library("tidyr")
library("reshape2")
library("gtable")
library("dplyr")
```

# Prediction with Covariates

```{r predict_with_covariates}
# Predicting at the BAUs when the large scale trend has covariates

# load meuse.grid and convert to SpatialPixelsDataFrame

data("meuse", package="sp")
coordinates(meuse) = ~ x + y
data("meuse.grid", package = "sp")
coordinates(meuse.grid) = ~ x + y
gridded(meuse.grid) = TRUE

# Remove any common fields in the data and the BAUs (recall that in FRK all covariates
# need to be with the BAUs)

meuse$soil <- meuse$dist <- meuse$ffreq <- NULL

# formula for SRE
f <- log(zinc) ~ 1 + sqrt(dist)

# Run FRK
S <- FRK(f = f,                # formula
         data = list(meuse),    # data (just meuse)
         BAUs = meuse.grid,     # the BAUs
         regular = 0)           # irregularly placed basis functions

# Predict at all the BAUs
Pred <- predict(S, obs_fs = FALSE)  
Pred_df<-SpatialPolygonsDataFrame_to_df(sp_polys = Pred,vars=c("mu","sd"))
```

```{r plot_predictions_1}
g1<-ggplot() +
  geom_polygon(aes(x=x, y=y, fill=mu,group=id),
               data=Pred_df)
g1<-g1 + xlab("Easting") + ylab("Northing")

g2<-ggplot() +
  geom_polygon(aes(x=x, y=y, fill=sd, group=id),
               data=Pred_df)
g2<-g2 + xlab("Easting") + ylab("Northing")

grid.arrange(g1,g2,nrow=1)
```

### Data and Prediction Locations in Same Data Frame (missing response data)

```{r plot_predictions_2}
# Example where both observed data and prediction locations are
# supplied in one data frame, with response at prediction locations
# missing (NA)

# Reload meuse data and assume first 10 observations are missing
data("meuse", package = "sp")
meuse[1:10, "zinc"] <- NA

# Create a complete-data data frame
meuse2 <- subset(meuse, !is.na(zinc))
meuse2 <- meuse2[, c("x", "y", "zinc")]
coordinates(meuse2) <- ~x+y

# Create BAUs around all prediction and observation 
# points (after removing data from field)

meuse$zinc <- NULL
coordinates(meuse) <- c("x", "y")
meuse.grid2 <- BAUs_from_points(meuse)

# Run the function FRK with the data frame and created BAUs.  
f <- log(zinc) ~ 1 + sqrt(dist)
S <- FRK(f = f, 
         data = list(meuse2), 
         BAUs = meuse.grid2, 
         regular = 0)
Pred <- predict(S, obs_fs = FALSE)
data(meuse)
plot(Pred$mu[1:10],log(meuse$zinc[1:10]),xlab="Predicted",ylab="Actual",cex.lab=1.5)
abline(c(0,1))
```

### Predictions of Spatial Averages

```{r predict_areal_support1}
# Construct model for meuse data with square grid, and consider
# prediction grid with different support

# Initialise seed to ensure reproducibility
set.seed(1)

# Load meuse data and convert to sp object
data("meuse", package = "sp")
coordinates(meuse) =  ~ x + y

# Construct the BAUs
GridBAUs1 <- auto_BAUs(manifold = plane(),    # we are on the plane
                       type = "grid",         # gridded BAUs (not hex)
                       cellsize = c(100, 100), # 100m x 100m
                       data = meuse,          # data for boundary
                       nonconvex_hull = TRUE, # nonconvex boundary
                       convex = -0.05)        # convex parameter (see
                                              # INLA::inla.nonconvex.hull)
```

```{r predict_areal_support2}
G <- auto_basis(manifold = plane(),  # we are on the plane
                data = meuse,        # data around which to construct basis
                regular = 0,         # irregularly placed basis
                nres = 3,            # three resolutions
                type = "bisquare")   # bisquare functions

## Plot basis functions
dev.new()
print(show_basis(G) +            # main call
      xlab("Easting (m)") +      # x-label
      ylab("Northing (m)") +     # y-label
      coord_fixed())             # fixed asp. ratio
```


```{r predict_areal_support3}
# formula (just intercept model)
f <- log(zinc) ~ 1


S <- FRK(f = f,                     
         data = list(meuse),        
         BAUs = GridBAUs1,          
         basis = G,                 
         average_in_BAU = FALSE)    
```


```{r predict_areal_support4}
# Now look at prediction of averages on a 600m x 600m grid
Pred_regions <- auto_BAUs(manifold = plane(),     # we are on the plane
                          cellsize = c(600, 600),  # 600m x 600m
                          type = "grid",          # grid
                          data = meuse,           # meuse data for boundary
                          convex = -0.05)         # convex parameter (see
                                                  # INLA::inla.nonconvex.hull)

## Now predict over these regions
Pred <- predict(S, newdata = Pred_regions,   # prediction regions
                obs_fs = FALSE)              # Case 2 in paper (default)

## Plot results

## First converg SpatialPolygonsDataFrame to data frame
Pred_regions_df <- SpatialPolygonsDataFrame_to_df(sp_polys = Pred,      # the sp obj
                                                  vars = c("mu", "sd")) # the variables to put
                                                                        # into the data frame
```

```{r predict_areal_support5}
g1<-ggplot() +
  geom_polygon(aes(x=x, y=y, fill=mu,group=id),
               data=Pred_regions_df)
g1<-g1 + xlab("Easting") + ylab("Northing")

g2<-ggplot() +
  geom_polygon(aes(x=x, y=y, fill=sd, group=id),
               data=Pred_regions_df)
g2<-g2 + xlab("Easting") + ylab("Northing")

grid.arrange(g1,g2,nrow=1)
```

### Data which are Spatial Averages (Areal Data)

```{r predict_areal_data1}
# Now consider an example where the observed data do not have
# point support

set.seed(1)      # Fix seed
data("meuse.grid", package = "sp") # load meuse data grid
data("meuse", package = "sp")      # load meuse data

## Generate observations with large spatial support
meuse_pols <- NULL  # initialise
offset <- 150       # half-size of box (150 m)

## for each meuse data point
for (i in 1:nrow(meuse)) {
    this_meuse <- meuse[i, ]
    meuse_pols <- rbind(meuse_pols,  # create a box aroun it of sides 300 m in length
                        data.frame(x = c(this_meuse$x - offset,
                                         this_meuse$x + offset,
                                         this_meuse$x + offset,
                                         this_meuse$x - offset),
                                   y = c(this_meuse$y - offset,
                                         this_meuse$y - offset,
                                         this_meuse$y + offset,
                                         this_meuse$y + offset),
                                   id = i,
                                   zinc = this_meuse$zinc))
}

## Convert to SpatialPolygonsDataFrame
meuse_pols <- df_to_SpatialPolygons(meuse_pols,coords = c("x", "y"), keys = "id", proj = CRS())
meuse_pols <- SpatialPolygonsDataFrame(meuse_pols,
                                       data.frame(row.names = row.names(meuse_pols),
                                                  zinc = meuse$zinc))
coordnames(meuse_pols) <- c("x", "y")

## Convert meuse and meuse.grid to sp object
coordinates(meuse) <- ~x + y
coordinates(meuse.grid) <- ~x+y

```

```{r conditional_simulation}
# Conditionally simulate field on the BAUs and aggregate to polygons 
# to generate realistic observations

# First construct the BAUs
GridBAUs <- auto_BAUs(manifold = plane(),      # 2D plane
                      cellsize = c(50, 50),    # BAU cellsize
                      type = "grid",           # grid (not hex)
                      data = meuse,            # data around which to create BAUs
                      convex = -0.05)          # border buffer factor
GridBAUs$fs <- 1

## Now fit variogram to meuse data and conditionally simulate
f <- log(zinc) ~ 1
lzn.vgm <- variogram(f, meuse)
lzn.fit <- fit.variogram(lzn.vgm, model = vgm(1, "Sph", 900, 1))
lzn.sim <- krige(formula = f, meuse,
                 GridBAUs,
                 model = lzn.fit,
                 nsim=1, nmax = 30)

# Observe field using large spatial support. In the below we run FRK to just generate
# the mapping matrix, which we then use to assign values to the meuse polygons
S <- FRK(f = f,  ## Just get out C Matrix
         data = list(meuse_pols),
         BAUs = GridBAUs,
         regular = 0,
         n_EM = 1, print_lik = FALSE)

meuse_pols$zinc <- as.numeric(S@Cmat %*% exp(as.numeric(lzn.sim$sim1)) +
                              0.1*rnorm(length(meuse_pols)))

```


```{r predict_areal_data2}
## Generate the basis functions
G <- auto_basis(manifold = plane(),   # 2D plane
                data = meuse,         # meuse data
                nres = 3,             # number of resolutions
                type = "bisquare",    # type of basis function
                regular = 0)          # place irregularly in domain

## Construct and fit the SRE model

S <- FRK(f = f,
         data=list(meuse_pols),
         BAUs=GridBAUs,
         basis=G)
         
GridBAUs2 <- predict(S, obs_fs = FALSE)
```

```{r predict_areal_data3}
## Create data frames for plotting
BAUs_df <- data.frame(GridBAUs2)
Obs_df <- SpatialPolygonsDataFrame_to_df(sp_polys = meuse_pols,   # BAUs to convert
                                         vars = c("zinc"))  # fields to extract

## Now plot the footprints
g1 <- ggplot() +
    geom_path(data = Obs_df,
              aes(x, y, group = id),
              colour = "black") +
    coord_fixed() +
    xlab("Easting (m)") + ylab("Northing (m)")


## Now plot the prediction
g2 <- ggplot() +                         # Use a plain theme
    geom_tile(data = BAUs_df,            # Draw BAUs
              aes(x, y, fill = mu),      # Colour <-> Mean
              colour = "light grey") +   # Border is light grey
    scale_fill_distiller(palette = "Spectral", name = "pred")  +  # Spectral palette
    coord_fixed() +                               # fix aspect ratio
    xlab("Easting (m)") + ylab("Northing (m)")    # axes labels

## Now plot the prediction error
g3 <- ggplot() +                          # Similar to above but with s.e.
    geom_tile(data = BAUs_df,
              aes(x, y, fill = sqrt(var)),
              colour = "light grey") +
    scale_fill_distiller(palette = "BrBG",
                         name = "s.e.") +
    coord_fixed() +
    xlab("Easting (m)") + ylab("Northing (m)")

dev.new()
plot(g1)

```

```{r}

#library(gtable)
#p2 <- ggplotGrob(g2)
#p3 <- ggplotGrob(g3)
#g <- cbind(p2, p3, size="first")
#g$heights <- unit.pmax(p2$heights, p3$heights)
#grid.newpage()
#grid.draw(g)

grid.arrange(g2, g3, nrow = 1) # Plot
```

```{r Poisson_example_BAUs}

x.seq <- y.seq <- seq(0, 1, length.out = 100)
BAUs <- expand.grid(x = x.seq, y = y.seq)
coordinates(BAUs) = ~ x + y
gridded(BAUs) <- TRUE

BAUs_df <- coordinates(BAUs) %>% as.data.frame

```

```{r Poisson_generate_data}

set.seed(2020)

# Define some complicated smooth trigonometric field

#f <- function(x, y, a = 0, b = 0, l = 1) exp(-l * sqrt((x - a)^2 + (y - b)^2))
logistic <- function(x, L = 1, k=1, x0 = 0) L / (1 + exp(-k * (x - x0)))
smooth_Y_process <- function(x, y) {
    a <- 4 + 2 * sin(5 * x) + 2 * cos(4 * y) + 2 * sin(3 * x * y)+
    sin(7 * x) + cos(9 * y) +
    sin(12 * x) + cos(17 * y) + sin(14 * x) * cos(16 * y)  +
    sin(23 * x) + cos(22 * y) + sin(24 * x) * cos(26 * y) +
    x + y + x^2 + y^2 + x * y + 
    sin(10 * pi * x * y) + cos(20 * x * y - 3) + sin(40 * x * y)
    
  logistic(a, x0 = 2, L = 6, k = 0.35)
}
  
BAUs_df <- BAUs_df %>% 
  dplyr::mutate(Y = smooth_Y_process(x, y) + rnorm(length(BAUs), mean = 0, sd = 0.2)) 

## Compute the true mean process, mu, over the BAUs
BAUs_df <- dplyr::mutate(BAUs_df, mu = exp(Y))

## Subsample n of the BAUs to act as observation locations, and simulate data
n <- 750
Poisson_simulated <- sample_n(BAUs_df, n) %>% 
  dplyr::mutate(Z = rpois(n, lambda = mu)) %>%
  dplyr::select(x, y, Z)  

## scalar matrix for fine scale variation, and converts to SpatialPixelsDF
BAUs$fs <- rep(1, length(BAUs)) 

## Convert Poisson_simulated to Spatial* object
coordinates(Poisson_simulated) <- ~ x + y

```

```{r Poisson_example_SRE}
S <- FRK(f = Z ~ 1, data = list(Poisson_simulated), 
                      nres = 3, BAUs = BAUs, 
                      response = "poisson", 
                      link = "log", 
                      K_type = "precision", method = "TMB", est_error = FALSE) 
pred_list<- predict(S, type = c("link", "mean"))

## Predictions, uncertainty, and data
plot_list <- plot(S, pred_list$newdata, 
                  labels_from_coordnames = FALSE)

```

```{r plot_BAUs}

g1<-ggplot(BAUs_df) + geom_tile(aes(x, y, fill = mu)) + 
  scale_fill_distiller(palette = "Spectral") + 
  labs(x = expression(s[1]), y = expression(s[2])) +
  theme_bw() + coord_fixed() + 
  scale_x_continuous(breaks=c(0.25, 0.75), expand = c(0, 0)) + 
  scale_y_continuous(breaks=c(0.25, 0.75), expand = c(0, 0)) + 
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 19), 
        legend.text = element_text(size = 16), 
        plot.title = element_text(hjust = 0.5, size = 19))
g1

```

```{r plot_data}

g2<-plot_list$Z + labs(title = plot_list$Z$labels$fill) + 
  labs(fill = "", colour = "") + 
  scale_x_continuous(breaks=c(0.25, 0.75), expand = c(0, 0)) + 
  scale_y_continuous(breaks=c(0.25, 0.75), expand = c(0, 0)) + 
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 19), 
        legend.text = element_text(size = 16), 
        plot.title = element_text(hjust = 0.5, size = 19))
g2

```

```{r plot_mu}

g3<-plot_list$p_Y + labs(title = plot_list$p_Y$labels$fill) + 
  labs(fill = "", colour = "") + 
  scale_x_continuous(breaks=c(0.25, 0.75), expand = c(0, 0)) + 
  scale_y_continuous(breaks=c(0.25, 0.75), expand = c(0, 0)) + 
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 19), 
        legend.text = element_text(size = 16), 
        plot.title = element_text(hjust = 0.5, size = 19))
g3

```




